Utilizing Parallel Workers: \\LLVM's Vectorization Plan

### 2024-11-20 paper/paper.tex
paper

@@

LLVM's VPlan is currently being integrated into the existing Loop Vectorizer and is
already modeling most inner loop vectorizations and transformations~\cite{llvmvplanupdate}. 
Vectorization for outer loops is also in development and can be enabled by setting 
the \texttt{-enable-vplan-native-path} flag~\cite{llvmouterloop}. In the future, the plan will 
be to merge both of these loop vectorization paths into one.

@@

\section{LLVM's Vectorization Plan}
\label{sec:vplan}
Lorem ipsum

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

LLVM's VPlan is currently being integrated into the existing Loop Vectorizer and is
already modeling most inner loop vectorizations and transformations~\cite{llvmvplanupdate}. 
Vectorization for outer loops is also in development and can be enabled by setting 
the \texttt{-enable-vplan-native-path} flag~\cite{llvmouterloop}. In the future, the plan will 
be to merge both of these loop vectorization paths into one. Merging the existing SLP Vectorizer
with VPlan is also planned.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{LLVM's Vectorization Plan}
\label{sec:vplan}
Originally LLVM implements two different vectorizers for auto-vectorization, LoopVectorize (LV)
and SLPVectorize. These vectorizers are a part of LLVM's middle-end and run after module 
simplification. As VPlan has not yet been integrated into SLP-Vectorization we will only focus on
explaining the original LoopVectorize structure.

LoopVectorize was designed to focus only on vectorization of innermost loops~\cite{llvmintrvplan}. 
When starting with a initial scalable loop in IR it will first run a legality phase to check 
whetever and how it is legal to vectorize the loop. This phase would produce artifacts like if 
and what runtime aliasing checks would be required. If some vectorization would be legal the 
cost model would run next, creating more artifacts based on the scalar IR and legality artifacts. 
These cost model artifacts would associate different vectorization possibilities with a fixed cost.
Lastly the transformation phase would vectorize the underlying IR based on all the produced 
artifacts.

However, this LoopVectorize architecture had several problems. For once extending the rigid system
to include outer loop and nested loop vectorizations would have been very challenging. Producing
and considering more and more artifacts when trying to transform the scalar into a vectorized
loop all at once does not scale well.

\subsection{State of Vectorization Plan (VPlan)}

\subsection{Future of VPlan}
%\subsection{Integration with LV}
%\subsection{Outer Loop Vectorization}
%\subsection{Merging with SLP Vectorize}
%\subsection{Function Vectorization}

\begin{figure}
  \centering
  \includegraphics[width=0.475\textwidth]{images/vplan-future.png}
  \caption{The planned future architecture for VPlan. The VPlan optimization would
  work as iteratievely applying VPlan-to-VPlan transformations.}
  \label{fig:vplan-future}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
